---
title: "VDJ data analysis on integrated dataset, Quantile: `r params$chosen.quantile`, Integration: `r params$integration.case`"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: no
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: true
    theme: lumen
params:
  outdir: NA
  PROJECT: NA
  config.version: NA
  output.version: NA
  chosen.quantile: NA
  integration.case: NA
---

```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

<style type="text/css">
    div.datatables { height: auto !important;}
</style>

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=14, fig.height=10}
#####----------------------------------------------------------------------#####
#
# trnguyen@ukaachen.de
#
#####----------------------------------------------------------------------#####
# gc()
# rm(list = ls())

scrna_pipeline_src <- "/media/hieunguyen/HNSD01/src/src_pipeline/scRNA_GEX_pipeline/processes_src"
source(file.path(scrna_pipeline_src, "import_libraries.R"))
source(file.path(scrna_pipeline_src, "helper_functions.R"))
source(file.path(scrna_pipeline_src, "s8_integration_and_clustering.R"))

path.to.project.src <- "/media/hieunguyen/HNSD01/src/bcr_data_analysis/240805_data_analysis"
source(file.path(path.to.project.src, "config.R"))
#####----------------------------------------------------------------------#####
# CONFIGURATIONS 
#####----------------------------------------------------------------------#####
analysis.round <- "1st"
chosen.seed <- 42
num.dim.integration <- 25 
num.PCA <- 25
num.dim.cluster <- 25
num.PC.used.in.Clustering <- 25
num.PC.used.in.UMAP <- 25
my_random_seed <- 42

# install.packages("https://cran.r-project.org/src/contrib/Archive/knitr/knitr_1.45.tar.gz", type = "source", repos = NULL)
# install.packages("https://cran.r-project.org/src/contrib/Archive/rmarkdown/rmarkdown_2.26.tar.gz", type = "source", repos = NULL)

all.samples <- c("M1", "M2", "M3", "P1", "P2", "P3")
all.quantiles <- c(0.99, 0.95, 0.90, 0.85)
all.integration.case <- list(
  all_samples = all.samples,
  mouse1 = c("M1", "P1"),
  mouse2 = c("M2", "P2"),
  mouse3 = c("M3", "P3")
)

#####----------------------------------------------------------------------#####
##### input arguments
#####----------------------------------------------------------------------#####

# outdir <- "/media/hieunguyen/HNSD01/outdir"
# PROJECT <- "240805_BSimons"
# output.version <- "20240820"
# config.version <- "default"
# chosen.quantile <- 0.85
# integration.case <- "all_samples"

outdir <- params$outdir
PROJECT <- params$PROJECT
output.version <- params$output.version
config.version <- params$config.version
chosen.quantile <- params$chosen.quantile
integration.case <- params$integration.case

path.to.main.input <- file.path(outdir, PROJECT, output.version, config.version)
path.to.main.output <- file.path(outdir, PROJECT, "data_analysis", output.version, config.version)

path.to.01.output <- file.path(path.to.main.output, "01_output")
path.to.04.output <- file.path(path.to.main.output, "04_output", sprintf("quantile_%s", chosen.quantile))
path.to.05.output <- file.path(path.to.main.output, "05_output", sprintf("quantile_%s", chosen.quantile), integration.case)
path.to.06.output <- file.path(path.to.main.output, "06_output", sprintf("quantile_%s", chosen.quantile), integration.case)
dir.create(path.to.06.output, showWarnings = FALSE, recursive = TRUE)

path.to.input.s.obj <- file.path(path.to.04.output, integration.case, "s8_output", sprintf("%s.output.s8.rds", PROJECT))
s.obj <- readRDS(path.to.input.s.obj)

meta.data <- s.obj@meta.data %>% rownames_to_column("barcode")
meta.data <- meta.data %>% rowwise() %>%
  mutate(SampleID = sprintf("%s_%s", name, HTO_classification)) %>%
  column_to_rownames("barcode")
meta.data <- meta.data[row.names(s.obj@meta.data),]
s.obj <- AddMetaData(object = s.obj, col.name = "SampleID", metadata = meta.data$SampleID)

#####----------------------------------------------------------------------#####
##### pre-processing the data with ibex
#####----------------------------------------------------------------------#####
#### install scRepertoire 2.0
if (packageVersion("scRepertoire") != "2.0.6"){
  install.packages("https://cran.r-project.org/src/contrib/Archive/Matrix/Matrix_1.6-5.tar.gz", type = "source", repos = NULL)
  install.packages(c("ggdendro", "iNEXT", "quantreg"))
  # install.packages("https://www.bioconductor.org/packages/release/bioc/src/contrib/scRepertoire_2.0.0.tar.gz", type = "source", repos = NULL)
  devtools::install_github(repo = "ncborcherding/scRepertoire", upgrade = FALSE)
  install.packages("https://cran.r-project.org/src/contrib/Archive/ggplot2/ggplot2_3.4.4.tar.gz", repos = NULL, type = "source")
}
library(scRepertoire)

library("Ibex")

threshold <- 0.85
removeNA <- TRUE
removeMulti <- TRUE
if (file.exists(file.path(path.to.06.output, sprintf("combined_contigs_%s_%s_%s.rds", threshold, removeNA, removeMulti))) == FALSE){
  path.to.storage <- "/media/hieunguyen/HNSD01/storage"
  path.to.main.input <- file.path(path.to.storage, PROJECT)
  
  path.to.VDJ.input <- file.path(path.to.main.input, "VDJ")
  
  all.contig.files <- Sys.glob(file.path(path.to.VDJ.input, "*/filtered_contig_annotations.csv"))
  
  sample.names <- unlist(lapply(lapply(all.contig.files, dirname), basename))
  
  names(all.contig.files) <- sample.names
  
  contig_list <- lapply(all.contig.files, vroom, show_col_type = FALSE)
  
  names(contig_list) <- sample.names
  combined.contigs <- combineBCR(contig_list,
                                   samples = sample.names,
                                   ID = sample.names,
                                   removeNA = removeNA,
                                   removeMulti = removeMulti,
                                   threshold = threshold)
  
  for (i in names(combined.contigs)){
    tmpdf <- combined.contigs[[i]]
    sample.id <- unique(tmpdf$ID)
    tmpdf <- tmpdf %>% rowwise() %>%
      mutate(barcode = str_replace(barcode, sprintf("%s_%s", sample.id, sample.id), sample.id))
    combined.contigs[[i]] <- tmpdf
  }
  
  saveRDS(combined.contigs, file.path(path.to.06.output, sprintf("combined_contigs_%s_%s_%s.rds", threshold, removeNA, removeMulti)))
} else {
  combined.contigs <- readRDS(file.path(path.to.06.output, sprintf("combined_contigs_%s_%s_%s.rds", threshold, removeNA, removeMulti)))
}

# 
# if (file.exists(file.path(path.to.06.output, sprintf("%s.ibex.output.s8.rds", PROJECT))) == FALSE){
#   ##### change the barcode
#     
#   s.obj <- combineExpression(combined.contigs, s.obj, cloneCall="aa")
#   
#   #Subset only B cells (by contigs)
#   s.obj$BCR.recoverd <- "No"
#   s.obj$BCR.recoverd[!is.na(s.obj$CTaa)] <- "Yes"
#   s.obj <- subset(s.obj, BCR.recoverd == "Yes")
#   
#   s.obj <- runIbex(s.obj, 
#                        chains = "Heavy",
#                        encoder.input = "KF", 
#                        reduction.name = "ibex.KF")
#   
#   s.obj <- RunUMAP(s.obj, 
#                    reduction = "ibex.KF",
#                    dims = 1:30,
#                    reduction.name = 'ibex.umap', 
#                    reduction.key = 'ibexUMAP_')
#   
#   saveRDS(s.obj, file.path(path.to.06.output, sprintf("%s.ibex.output.s8.rds", PROJECT)))
# } else {
#   s.obj <- readRDS(file.path(path.to.06.output, sprintf("%s.ibex.output.s8.rds", PROJECT)))
# }
```

# UMAP
## Cells with proper clone information and cells without proper clone information
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
# DimPlot(object = s.obj, cells.highlight = list("with.Clone" = with.clone.cells, "without.Clone" = without.clone.cells), label = TRUE, label.size = 8, reduction = "INTE_UMAP")
```

## UMAP: All cells in all samples
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
DimPlot(object = s.obj, label = TRUE, label.size = 8, reduction = "INTE_UMAP", label.box = TRUE)
```

## UMAP: all cells, grouped by samples
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
DimPlot(object = s.obj, label = TRUE, label.size = 8, reduction = "INTE_UMAP", label.box = TRUE, group.by = "name")
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
s.obj <- combineExpression(combined.contigs, s.obj, cloneCall="aa")

all.available.ht <- list()
for (sample.id in unique(s.obj$name)){
  all.available.ht[[sample.id]] <- unique(subset(s.obj, name == sample.id)$HTO_classification)  %>% sort()
}

if (file.exists(file.path(path.to.06.output, sprintf("clone_Table_%s.xlsx", integration.case))) == FALSE){
  meta.data <- s.obj@meta.data
  count.clones <- table(meta.data$CTstrict)
  
  clonedf <- data.frame(CTstrict = setdiff(unique(meta.data$CTstrict), c(NA)))
  clonedf <- clonedf %>% rowwise() %>%
    mutate(count = count.clones[[CTstrict]]) %>% arrange(desc(count))
  
  for (sample.id in unique(s.obj$name)){
    print(sprintf("working on sample %s", sample.id))
    clonedf[[sample.id]] <- unlist(lapply(clonedf$CTstrict, function(x){
      nrow(subset(meta.data, meta.data$CTstrict == x & meta.data$name == sample.id))
    }))  
    for (ht in all.available.ht[[sample.id]]){
      print(sprintf("working on sample %s hashtag %s", sample.id, ht))
      clonedf[[sprintf("%s_%s", sample.id, ht)]] <- unlist(lapply(clonedf$CTstrict, function(x){
        nrow(subset(meta.data, meta.data$CTstrict == x & meta.data$name == sample.id &  meta.data$HTO_classification == ht))
      })) 
    }
  }
  writexl::write_xlsx(clonedf, file.path(path.to.06.output, sprintf("clone_Table_%s.xlsx", integration.case)))
} else {
  clonedf <- readxl::read_excel(file.path(path.to.06.output, sprintf("clone_Table_%s.xlsx", integration.case)))
}
```

## UMAP clone size
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
DimPlot(object = s.obj, reduction = "INTE_UMAP", group.by = "cloneSize")
```

# Table: All clones in the dataset
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
clonedf %>% create_dt()
```

# Plot clones on UMAP
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
top6.clones <- hash()
for (i in setdiff(colnames(clonedf), c("CTstrict", "count"))){
  top6.clones[[i]] <- clonedf[c("CTstrict", i)]
  colnames(top6.clones[[i]]) <- c("CTstrict", "count")
  top6.clones[[i]] <- top6.clones[[i]] %>% arrange(desc(count)) %>% head(6) %>% pull(CTstrict)
}
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
for (sample.id in all.integration.case[[integration.case]]){
  cat(sprintf("## Sample %s {.tabset} \n", sample.id))
  tmp.s.obj <- subset(s.obj, name == sample.id)
  tmp.metadata <- tmp.s.obj@meta.data
  clone.barcodes <- lapply(top6.clones[[sample.id]], function(x){
    subset(tmp.metadata, CTstrict == x) %>% rownames()
  })
  names(clone.barcodes) <- top6.clones[[sample.id]]
  for (clone in names(clone.barcodes)){
    cat(sprintf("### Clone %s \n", clone))
    p <- DimPlot(object = tmp.s.obj, reduction = "INTE_UMAP", cells.highlight = clone.barcodes[[clone]]) +
      theme(legend.position = "none")
    print(p)
    cat("\n \n")  
  }
  cat("\n \n")
}
```

# Basic clonal analysis

## Clonal Quant
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
names(combined.contigs) <- to_vec(for (item in names(combined.contigs)) str_split(item, "_")[[1]][[1]] )
meta.data <- s.obj@meta.data

if (file.exists( file.path(path.to.06.output, "combined_contigs_with_hashtags_info.rds")) == FALSE){
  new.combined.contigs <- list()
  for (sample.id in names(combined.contigs)){
    tmp.combined.contigs <- combined.contigs[[sample.id]]
    for (ht in unique(subset(meta.data, meta.data$name == sample.id)$HTO_classification)){
      print(sprintf("working on sample %s, hashtag %s", sample.id, ht))
      ht.barcodes <- subset(meta.data, meta.data$name == sample.id & meta.data$HTO_classification == ht) %>% row.names()
      new.combined.contigs[[sprintf("%s_%s", sample.id, ht)]] <- subset(tmp.combined.contigs, tmp.combined.contigs$barcode %in% ht.barcodes)
      print(dim(new.combined.contigs[[sprintf("%s_%s", sample.id, ht)]]))
    }
  }
  saveRDS(new.combined.contigs, file.path(path.to.06.output, "combined_contigs_with_hashtags_info.rds"))  
} else {
  new.combined.contigs <- readRDS( file.path(path.to.06.output, "combined_contigs_with_hashtags_info.rds"))
}
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
clonalAbundance(new.combined.contigs, 
                cloneCall = "gene", 
                scale = FALSE)
```

## Clonal compare between hashtags in each sample {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
for (sample.id in all.integration.case[[integration.case]]){
  cat(sprintf("### Sample %s \n", sample.id))
  p <- clonalCompare(new.combined.contigs, 
                  top.clones = 10, 
                  samples = to_vec(for (item in names(new.combined.contigs)) if(grepl(sample.id, item)) item), 
                  cloneCall="CTstrict", 
                  graph = "alluvial")
  print(p)
  cat("\n \n")
}
```

## Clonal compare between paired samples 

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
cat(sprintf("### Alluvial plot between samples %s", paste(all.integration.case[[integration.case]], collapse = ", ")))
p <- clonalCompare(combined.contigs, 
                  top.clones = 10, 
                  samples = all.integration.case[[integration.case]], 
                  cloneCall="CTstrict", 
                  graph = "alluvial")
print(p)
```

# Compare to clones from the bulk-BCR data
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
#####----------------------------------------------------------------------------#####
##### PREPROCESS CLONE INFORMATION FROM BULK DATA AND SINGLE CELL DATA; MATCH TWO DATASET TOGETHER.
#####----------------------------------------------------------------------------#####
if (file.exists(file.path(path.to.06.output, "combine_clonedf.xlsx")) == FALSE){
  path.to.bulk.data <- file.path(outdir, "240826_BSimons", "data_analysis")
  path.to.04.bulk.output <- file.path(path.to.bulk.data, "04_output")
  path.to.02.bulk.output <- file.path(path.to.bulk.data, "02_output")
  
  if (file.exists(file.path(path.to.06.output, "sc_clones.xlsx")) == FALSE){
    all.clone.files <- Sys.glob(file.path(path.to.04.bulk.output, "CDR3_0.15", "clonesets_*.xlsx"))
    all.clone.files <- all.clone.files[grepl("split", all.clone.files) == FALSE]
    
    names(all.clone.files) <- to_vec(for (item in basename(all.clone.files)) str_replace(str_replace(item, "clonesets_", ""), ".xlsx", ""))
    
    bulkdf <- data.frame()
    for (i in names(all.clone.files)){
      file <- all.clone.files[[i]]
      tmpdf <- readxl::read_excel(file)
      tmpdf$MID <- i 
      bulkdf <- rbind(bulkdf, tmpdf)
    }
    bulk.selected.cols <- c("V.gene", "J.gene", "nSeqCDR3", "aaSeqCDR3", "MID", "name")
    bulkdf.raw <- bulkdf
    bulkdf$V.gene <- unlist(lapply(bulkdf$V.gene, function(x){
      str_split(x, "[*]")[[1]][[1]]
    }))
    bulkdf$J.gene <- unlist(lapply(bulkdf$J.gene, function(x){
      str_split(x, "[*]")[[1]][[1]]
    }))
    bulkdf$name <- bulkdf$MID
    bulkdf <- bulkdf[, bulk.selected.cols]
    path.to.sc.data <- file.path(outdir, "240805_BSimons", "20240820", "default", "VDJ_output")
    all.vdj.outputs <- Sys.glob(file.path(path.to.sc.data, "annotated_contigs_clonaltype*.csv"))
    
    scdf <- data.frame()
    for (file in all.vdj.outputs){
      sample.id <- str_replace(str_replace(basename(file), "annotated_contigs_clonaltype_", ""), ".csv", "")
      tmpdf <- read.csv(file)
      tmpdf$name <- sample.id
      scdf <- rbind(scdf, tmpdf)
    }
    scdf.raw <- scdf
    
    scdf <- subset(scdf, select = -c(X))
    selected.cols <- c("barcode", "name", "IGH", "cdr3_aa1", "cdr3_nt1")
    scdf <- scdf[, selected.cols]
    
    # remove NA clones
    scdf <- subset(scdf, is.na(scdf$IGH) == FALSE)
    
    ##### convert scdf to match bulkdf
    scdf$MID <- unlist(lapply(
      scdf$name, function(x){
        str_replace(str_replace(x, "M", "m"), "P", "m")
      }
    ))
    scdf$nSeqCDR3 <- scdf$cdr3_nt1
    scdf$aaSeqCDR3 <- scdf$cdr3_aa1
    
    scdf$V.gene <- unlist(lapply(scdf$IGH, function(x){
        str_split(x, "[.]")[[1]][[1]]
    }))
    scdf$J.gene <- unlist(lapply(scdf$IGH, function(x){
        str_split(x, "[.]")[[1]][[3]]
    }))
    
    scdf <- scdf[, c("barcode", bulk.selected.cols)]
    
    bulkdf$barcode <- to_vec( for(item in seq(1, nrow(bulkdf))) sprintf("seq%s", item) )
    writexl::write_xlsx(scdf, file.path(path.to.06.output, "sc_clones.xlsx"))
    writexl::write_xlsx(bulkdf, file.path(path.to.06.output, "bulk_clones.xlsx"))
  } else {
    scdf <- readxl::read_excel(file.path(path.to.06.output, "sc_clones.xlsx"))
    bulkdf <- readxl::read_excel(file.path(path.to.06.output, "bulk_clones.xlsx"))
  }
  
  scdf$origin <- "single-cell"
  bulkdf$origin <- "bulk"
  
  clonedf <- rbind(scdf, bulkdf[, colnames(scdf)])
  writexl::write_xlsx(clonedf, file.path(path.to.06.output, "combine_clonedf.xlsx"))
} else {
  clonedf <- readxl::read_excel(file.path(path.to.06.output, "combine_clonedf.xlsx"))
  scdf <- readxl::read_excel(file.path(path.to.06.output, "sc_clones.xlsx"))
  bulkdf <- readxl::read_excel(file.path(path.to.06.output, "bulk_clones.xlsx"))
}

##### add some more information on the sequences
clonedf <- clonedf %>% rowwise() %>% 
  mutate(VJ.len.combi = sprintf("%s_%s_%s", V.gene, J.gene,  nchar(nSeqCDR3)) )
```

# Group clone by sequence similarity and V-J genes
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
source(file.path(path.to.project.src, "helper_functions.R"))
clonedf <- clonedf %>% rowwise() %>% 
  mutate(VJ.len.combi = sprintf("%s_%s_%s", V.gene, J.gene,  nchar(nSeqCDR3)) )

##### Assign clones to clusters
if (file.exists(file.path(path.to.06.output, "full_clonedf.xlsx")) == FALSE){
  new.clonedf <- data.frame()
for (input.VJ.combi in unique(clonedf$VJ.len.combi)){
  tmpdf <- subset(clonedf, clonedf$VJ.len.combi == input.VJ.combi)
  seqs <- unique(tmpdf$aaSeqCDR3)
  if (length(seqs) >= 2){
    cluster.output <- assign_clusters_to_sequences(seqs, threshold = threshold)$res
    tmpdf[[sprintf("VJcombi_CDR3_%s", threshold)]] <- unlist(lapply(
      tmpdf$aaSeqCDR3, function(x){
        return(sprintf("%s_%s", input.VJ.combi, subset(cluster.output, cluster.output$seq == x)$cluster))
      }
    ))    
  } else {
    tmpdf[[sprintf("VJcombi_CDR3_%s", threshold)]] <- sprintf("%s_1", input.VJ.combi)
  }
  new.clonedf <- rbind(new.clonedf, tmpdf)
}
writexl::write_xlsx(new.clonedf, file.path(path.to.06.output, "full_clonedf.xlsx"))
} else {
  new.clonedf <- readxl::read_excel(file.path(path.to.06.output, "full_clonedf.xlsx"))
}

cloneCluster.summary <- data.frame(clone = unique(new.clonedf$VJcombi_CDR3_0.85)) %>% 
  rowwise() %>%
  mutate(num_samples = length(unique(subset(new.clonedf, new.clonedf$VJcombi_CDR3_0.85 == clone)$name))) %>%
  mutate(num_total_seq = nrow(subset(new.clonedf, new.clonedf$VJcombi_CDR3_0.85 == clone))) %>%
  mutate(num_unique_seq = length(unique(subset(new.clonedf, new.clonedf$VJcombi_CDR3_0.85 == clone)$nSeqCDR3))) %>%
  arrange(desc(num_total_seq))
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
print(sprintf("Number of clones: %s", nrow(cloneCluster.summary)))
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
print(sprintf("Number of clones that are unique in only ONE sample: %s", nrow(subset(cloneCluster.summary, cloneCluster.summary$num_samples == 1))))
```

# Visualization
## Circos plot
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
clone.sampledf <- data.frame(SampleID = unique(new.clonedf$name))
for (sample2 in unique(new.clonedf$name)){
  clone.sampledf[[sample2]] <- unlist(
    lapply(
      clone.sampledf$SampleID, function(sample1){
        return(length(unique(intersect(subset(new.clonedf, new.clonedf$name == sample1)$VJcombi_CDR3_0.85, subset(new.clonedf, new.clonedf$name == sample2)$VJcombi_CDR3_0.85))))
      }
    )
  )
}

clone.sampledf.pivot <- clone.sampledf %>% pivot_longer(!SampleID, names_to = "SampleID2", values_to = "count")

clone.sampledf %>% create_dt()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
if ("heatmaply" %in% installed.packages() == FALSE){
  install.packages("heatmaply")
}
library(heatmaply)
heatmaply(clone.sampledf %>% column_to_rownames("SampleID"))
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
chordDiagram(clone.sampledf.pivot) 
```

