---
title: "Data analysis second dataset - GEX, 26.10.2023"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    keep_md: true
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: false
    theme: lumen
---

```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

<style type="text/css">
    div.datatables { height: auto !important;}
</style>


```{r echo=FALSE, warning=FALSE, results='hide', message=FALSE}
# clean up 
gc()
rm(list = ls())

#####----------------------------------------------------------------------#####
# CONFIGURATIONS AND PREPARATION
#####----------------------------------------------------------------------#####
analysis.case <- "2nd_dataset_removed_5_6.without_reInt.res1"

# scrna_pipeline_src <- "/home/uk104163/src_2023/src_pipeline/scRNA_GEX_pipeline/processes_src"
scrna_pipeline_src <- "/home/hieunguyen/CRC1382/src_2023/src_pipeline/scRNA_GEX_pipeline/processes_src"
source(file.path(scrna_pipeline_src, "import_libraries.R"))
source(file.path(scrna_pipeline_src, "helper_functions.R"))

# path.to.project.src <- "/home/uk104163/src_2023/BSimons/BSimons_thesis_data"
path.to.project.src <- "/home/hieunguyen/CRC1382/src_2023/BSimons/CRC1382_BSimons_project/BSimons_scRNAseq_analysis"
source(file.path(path.to.project.src, "00_helper_functions.R"))

outdir <- "/home/hieunguyen/CRC1382/outdir"

path.to.output <- file.path(outdir, "BSimons", "OUTPUT", "THESIS_OUTPUT_20231026")

path.to.01.output <- file.path(path.to.output, "01_output")
path.to.02.output <- file.path(path.to.output, "02_output", analysis.case)
path.to.03.output <- file.path(path.to.output, "03_output", analysis.case)
dir.create(path.to.03.output, showWarnings = FALSE, recursive = TRUE)

path.to.save.table <- file.path(path.to.03.output, "tables")
path.to.save.figures <- file.path(path.to.03.output, "figures")

dir.create(path.to.save.table, showWarnings = FALSE, recursive = TRUE)
dir.create(path.to.save.figures, showWarnings = FALSE, recursive = TRUE)

s.obj <- readRDS(file.path(path.to.01.output, "2nd_dataset_removed_5_6.without_reInt.res1", "2nd_dataset_removed_5_6.without_reInt.res1.rds"))

count.YFP <- GetAssayData(s.obj, slot = "counts")["YFP", ]
count.YFP <- count.YFP[count.YFP != 0]
yfp.cells <- names(count.YFP)
non.yfp.cells <- to_vec(for (item in colnames(s.obj)) if (item %in% yfp.cells == FALSE) item)

s.obj.yfp <- subset(s.obj, cells = yfp.cells)
s.obj.nonyfp <- subset(s.obj, cells = non.yfp.cells)
```

# Some QC metrics after filtering out cells
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
  all.QC <- list()
  
  # Number of cells obtained per experiment/sample
  all.QC$cell.counts.plot <- ggplot(s.obj@meta.data, 
                                    aes(x=name , fill=name)) + 
    geom_bar() +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    theme(plot.title = element_text(hjust=0.5, face="bold", size = 14)) +
    ggtitle("Number of cells in each dataset")
  
  # distribution of number of UMI in each sample
  all.QC$nCountRNA.distribution <- ggplot(s.obj@meta.data,
                                          aes(color=name, x=nCount_RNA, fill = name)) + 
    geom_density(alpha = 0.2) +
    scale_x_log10() +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    theme(plot.title = element_text(hjust=0.5, face="bold", size = 14)) +
    ylab("Cell density") +
    geom_vline(xintercept = 500, color = "red") +
    ggtitle("Distribution of read depths in each sample")
  
  # distribution of number of features (genes detected) in each sample
  all.QC$nFeature_RNA.distribution <- ggplot(s.obj@meta.data,
                                             aes(color=name, x=nFeature_RNA, fill = name, y = ..scaled..)) + 
    geom_density(alpha = 0.2) +
    scale_x_log10() +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    theme(plot.title = element_text(hjust=0.5, face="bold", size = 14)) +
    ylab("Cell density") +
    geom_vline(xintercept = 1000, color = "red") +
    xlim(1000, 10000) +
    ggtitle("Distribution of number of detected genes in each sample")
  
  
  # scatter plot showing the relation between cell read-depth and number of genes detected.
  ## with Mitochondrial percentage
  all.QC$nCount.vs.nFeature.MT <- ggplot(s.obj@meta.data, 
                                         aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) + 
    geom_point() + 
    scale_colour_gradient(low = "gray90", high = "black") +
    stat_smooth(method=lm, formula = y ~ x) + # apply a linear regression to show the relation, if existed.
    scale_x_log10() + 
    scale_y_log10() + 
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    theme(plot.title = element_text(hjust=0.5, face="bold", size = 12)) +
    facet_wrap(~name) +
    ggtitle("Scatter plot: nCount_RNA vs. nFeature_RNA, cmap % Mitochondrial genes")
  
  ## with Ribosome percentage
  all.QC$nCount.vs.nFeature.Ribo <- ggplot(s.obj@meta.data, 
                                           aes(x=nCount_RNA, y=nFeature_RNA, color=percent.ribo)) + 
    geom_point() + 
    scale_colour_gradient(low = "gray90", high = "black") +
    stat_smooth(method=lm, formula = y ~ x) + # apply a linear regression to show the relation, if existed.
    scale_x_log10() + 
    scale_y_log10() + 
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    theme(plot.title = element_text(hjust=0.5, face="bold", size = 12)) +
    facet_wrap(~name) +
    ggtitle("Scatter plot: nCount_RNA vs. nFeature_RNA, cmap % Ribosome genes")
  
  s.obj@meta.data <- s.obj@meta.data %>% 
    mutate(log10GenesPerUMI = log10(nFeature_RNA) / log10(nCount_RNA))
  
  all.QC$complexity <- ggplot(s.obj@meta.data,
                              aes(x=log10GenesPerUMI, color = name, fill=name)) +
    geom_density(alpha = 0.2) +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    theme(plot.title = element_text(hjust=0.5, face="bold", size = 14)) +
    geom_vline(xintercept = 0.8) +
    ggtitle("Complexity: Log10(nCount_RNA) / log10(nFeature_RNA)")
  
  # add new slot for all.QC into the existed SEURAT OBJECT. 
  s.obj@misc$all.QC.after.filtering <- all.QC
```

## Raw data Quality control  {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
for (plot.name in names(s.obj@misc$all.QC)){
 
    cat('###',plot.name,'{.unlisted .unnumbered} \n')
    
    ##### 
    # plots or tables that we want to show in tabs
    #####
    print(s.obj@misc$all.QC[plot.name])
    cat(' \n \n')
}
```



## Ambient RNA background correction 

### DecontX clusters {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
for (plot.name in names(s.obj@misc$ambient.cluster.RNA.plot)){
 
    cat('####',plot.name,'{.unlisted .unnumbered} \n')
    
    ##### 
    # plots or tables that we want to show in tabs
    #####
    print(s.obj@misc$ambient.cluster.RNA.plot[plot.name])
    cat(' \n \n')
}
```

### Contamination level in each sample

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
ggplot(s.obj@meta.data, aes(x=name, y=AmbientRNA)) + 
  geom_boxplot()

```


### Ambient RNA contamination level {.tabset}

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
for (plot.name in names(s.obj@misc$ambient.contamination.plot)){
 
    cat('####',plot.name,'{.unlisted .unnumbered} \n')
    
    ##### 
    # plots or tables that we want to show in tabs
    #####
    print(s.obj@misc$ambient.contamination.plot[plot.name])
    cat(' \n \n')
}
```


## Descriptive statistics and filtering threshold {.tabset}
This section is devoted to the descriptive statistics of the following varialbes: `nFeature_RNA, nCount_RNA, percent.mt, percent.ribo`. 

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
for (plot.item in c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.ribo")){
  boxplot <- s.obj@meta.data %>% ggplot(aes_string(x = "name", y = plot.item)) +
    geom_boxplot(outlier.colour="black", outlier.shape=16,
               outlier.size=2, notch=FALSE) +
    ggtitle(sprintf("Boxplot: Distribution of %s in each dataset", plot.item))
  cat('###', plot.item,'{.unlisted .unnumbered} \n')
    
    ##### 
    # plots or tables that we want to show in tabs
    #####
  
  print(boxplot)
  
  cat(' \n \n')
}
```



## Descriptive statistics + UMAP {.tabset}

### % Mitochondrial
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
FeaturePlot(object = s.obj, reduction = "INTE_UMAP", feature = "percent.mt", label = TRUE, label.size = 8, pt.size = 0.5, label.color = "red", )
```

### % Ribosome
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
FeaturePlot(object = s.obj, reduction = "INTE_UMAP", feature = "percent.ribo", label = TRUE, label.size = 8, pt.size = 0.5, label.color = "red", )
```

### % nCount RNA
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
FeaturePlot(object = s.obj, reduction = "INTE_UMAP", feature = "nCount_RNA", label = TRUE, label.size = 8, pt.size = 0.5, label.color = "red", )
```

### % nGenes 
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
FeaturePlot(object = s.obj, reduction = "INTE_UMAP", feature = "nFeature_RNA", label = TRUE, label.size = 8, pt.size = 0.5, label.color = "red", )
```

## Number of cells in each cluster
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
count.cell.clusters <- table(s.obj@meta.data$seurat_clusters)  %>% as.data.frame()
colnames(count.cell.clusters) <- c("cluster", "count")

count.cell.clusters %>% create_dt()
```

## Number of cells in each colonization
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=15, fig.height=10}
s.obj@meta.data %>% subset(select = c(seurat_clusters, stage)) %>% table() %>% as.data.frame() %>% pivot_wider(names_from = "stage", values_from = "Freq") %>% create_dt()
```

## Cell cycle analysis
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
all.genes <- rownames(x = s.obj)
s.genes <- paste0("^", cc.genes$s.genes, "$", collapse = "|")
s.genes <- all.genes[grepl(s.genes, all.genes, ignore.case = TRUE)]
g2m.genes <- paste0("^", cc.genes$g2m.genes, "$", collapse = "|")
g2m.genes <- all.genes[grepl(g2m.genes, all.genes, ignore.case = TRUE)]

s.obj <- CellCycleScoring(s.obj, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
```

### UMAP grouped by cell cycle phase
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
DimPlot(object= s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, group.by = "Phase")
```

### PCA grouped by cell cycle phase
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
DimPlot(object= s.obj, reduction = "INTE_PCA", label = TRUE, label.box = TRUE, repel = TRUE, group.by = "Phase")
```


```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
Idents(s.obj) <- "seurat_clusters"
```

# Summary QC 
## Bioinformatics pipeline
Count, barcode, feature matrices were generated by CellRanger pipeline (v. 7.1.0) with default parameters. Sequence alignment and transcript quantification were done with a modified Mm10 reference genome, in which we manually added the gene YFP. Downstream analysis was performed by our custom pipeline written in R using Seurat v.4.0. Contaimnated ambient RNA was estimated by decontX (celda v1.12, https://genomebiology.biomedcentral.com/articles/10.1186/s13059-020-1950-6). Doublets were detected by DoubletFinder (https://www.sciencedirect.com/science/article/pii/S2405471219300730). Dimensional reductions, PCA and UMAP, as well as cells clustering, were done by built-in functions in Seurat. Samples were integrated by built-in SEURAT integration algorithm to remove any potential batch effects. 

Cells whose number of UMI less than 100 or number of genes less than 50 were filtered out. Moreover, cells having percentage of mitochondrial higher than 10% were also excluded from the analysis. By differential gene expression analysis between clusters versus clusters, cells populations were identified. We excluded some clusters and kept only populations of interest. 

## Summary QC table
We summarize the QC metrics for each sample in this table
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
##### HELPER FUNCTION TO GENERATE SUMMARY QC METRICS
generate_summary_qc <- function(meta.data){
  sumqcdf <- data.frame()
for (sample.id in unique(meta.data$name)){
  subset.metadata <- subset(meta.data, meta.data$name == sample.id)
  
  tmpdf <- data.frame(Sample = c(sample.id))
  tmpdf$min.nCount_RNA <- min(subset.metadata$nCount_RNA)
  tmpdf$max.nCount_RNA <- max(subset.metadata$nCount_RNA)
  tmpdf$mean.nCount_RNA <- mean(subset.metadata$nCount_RNA)
  tmpdf$median.nCount_RNA <- median(subset.metadata$nCount_RNA)
  
  tmpdf$min.nFeature_RNA <- min(subset.metadata$nFeature_RNA)
  tmpdf$max.nFeature_RNA <- max(subset.metadata$nFeature_RNA)
  tmpdf$mean.nFeature_RNA <- mean(subset.metadata$nFeature_RNA)
  tmpdf$median.nFeature_RNA <- median(subset.metadata$nFeature_RNA)
  
  tmpdf$min.percent.mt <- min(subset.metadata$percent.mt)
  tmpdf$max.percent.mt <- max(subset.metadata$percent.mt)
  tmpdf$mean.percent.mt <- mean(subset.metadata$percent.mt)
  tmpdf$median.percent.mt <- median(subset.metadata$percent.mt)
  
  tmpdf$min.percent.ribo <- min(subset.metadata$percent.ribo)
  tmpdf$max.percent.ribo <- max(subset.metadata$percent.ribo)
  tmpdf$mean.percent.ribo <- mean(subset.metadata$percent.ribo)
  tmpdf$median.percent.ribo <- median(subset.metadata$percent.ribo)
  
  tmpdf$min.log10GenesPerUMI <- min(subset.metadata$log10GenesPerUMI)
  tmpdf$max.log10GenesPerUMI <- max(subset.metadata$log10GenesPerUMI)
  tmpdf$mean.log10GenesPerUMI <- mean(subset.metadata$log10GenesPerUMI)
  tmpdf$median.log10GenesPerUMI <- median(subset.metadata$log10GenesPerUMI)
  
  tmpdf$min.AmbientRNA <- min(subset.metadata$AmbientRNA)
  tmpdf$max.AmbientRNA <- max(subset.metadata$AmbientRNA)
  tmpdf$mean.AmbientRNA <- mean(subset.metadata$AmbientRNA)
  tmpdf$median.AmbientRNA <- median(subset.metadata$AmbientRNA)
  
  tmpdf$num.Clones <- length(unique(subset.metadata$CTstrict))
  tmpdf$num.Clusters <- length(unique(subset.metadata$seurat_clusters))
  
  sumqcdf <- rbind(sumqcdf, tmpdf)
}
  return(sumqcdf)
}
```

### Raw data QC metrics
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
s.obj.raw <- readRDS(file.path(path.to.01.output, "raw_seurat_objects", "merge_2nd_datasets.raw.rds"))
meta.data.raw <- s.obj.raw@meta.data
sumqc.df.raw <- generate_summary_qc(meta.data.raw)
```


```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
sumqc.df.raw %>% create_dt()
```

### After filtering 

This is the summary QC metrics for the data we used in downstream analysis

```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
meta.data <- s.obj@meta.data
sumqcdf <- generate_summary_qc(meta.data)
```


```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
sumqcdf %>% create_dt()
```

# UMAP

## UMAP group by clusters
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.height=10, fig.width=15}
p <- DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE)
ggsave(plot = p, path = path.to.save.figures, filename = "UMAP_all_clusters_in_2nd_dataset.svg", device = "svg", dpi = 300, width = 10, height = 15)
print(p)
```

## UMAP group by samples
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.height=10, fig.width=15}
p <- DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, group.by = "name")
ggsave(plot = p, path = path.to.save.figures, filename = "UMAP_all_clusters_in_2nd_dataset.bySample.svg", device = "svg", dpi = 300, width = 10, height = 15)
print(p)
```

## UMAP: highlighted YFP+ cells
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
p <- DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, cells.highlight = yfp.cells, cols.highlight = "red") +
  theme(legend.position = "none")
ggsave(plot = p, path = path.to.save.figures, filename = "UMAP_all_clusters_highlighted_YFP_pos.svg", device = "svg", dpi = 300, width = 10, height = 15) 
print(p)
```


## UMAP: highlighted YFP+ cells, sample 132
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
p <- DimPlot(object = subset(s.obj, name == "Sample_132"), reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, cells.highlight = yfp.cells, cols.highlight = "red") +
  theme(legend.position = "none")
ggsave(plot = p, path = path.to.save.figures, filename = "UMAP_all_clusters_highlighted_YFP_pos.Sample132.svg", device = "svg", dpi = 300, width = 10, height = 15) 
print(p)
```

## UMAP: highlighted YFP+ cells, sample 133
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
p <- DimPlot(object = subset(s.obj, name == "Sample_133"), reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, cells.highlight = yfp.cells, cols.highlight = "red") +
  theme(legend.position = "none")
ggsave(plot = p, path = path.to.save.figures, filename = "UMAP_all_clusters_highlighted_YFP_pos.Sample133.svg", device = "svg", dpi = 300, width = 10, height = 15) 
print(p)
```

## UMAP: highlighted cells with clonal information, Sample 132
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
tmp.s.obj <- subset(s.obj, name == "Sample_132")
withVDJ.cells <- rownames(subset(tmp.s.obj@meta.data, is.na(tmp.s.obj@meta.data$CTstrict) == FALSE))
p <- DimPlot(object = tmp.s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, cells.highlight = withVDJ.cells, cols.highlight = "red") +
  theme(legend.position = "none")
ggsave(plot = p, path = path.to.save.figures, filename = "UMAP_all_clusters_highlighted_cells_with_clonalInfo.Sample132.svg", device = "svg", dpi = 300, width = 10, height = 15) 
print(p)
```

## UMAP: highlighted cells with clonal information, Sample 133
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
tmp.s.obj <- subset(s.obj, name == "Sample_133")
withVDJ.cells <- rownames(subset(tmp.s.obj@meta.data, is.na(tmp.s.obj@meta.data$CTstrict) == FALSE))
p <- DimPlot(object = tmp.s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, cells.highlight = withVDJ.cells, cols.highlight = "red") +
  theme(legend.position = "none")
ggsave(plot = p, path = path.to.save.figures, filename = "UMAP_all_clusters_highlighted_cells_with_clonalInfo.Sample133.svg", device = "svg", dpi = 300, width = 10, height = 15) 
print(p)
```

# Differential gene expression analysis between clusters
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
if (file.exists(file.path(path.to.03.output, "cluster_markers.rds")) == FALSE){
  cluster.markers <- FindAllMarkers(object = s.obj, assay = "RNA", test.use = "wilcox")
  saveRDS(cluster.markers, file.path(path.to.03.output, "cluster_markers.raw.rds"))
  cluster.markers <- subset(cluster.markers, cluster.markers$p_val_adj <= 0.05 & cluster.markers$avg_log2FC >= 0)
  saveRDS(cluster.markers, file.path(path.to.03.output, "cluster_markers.rds"))  
} else {
  cluster.markers <- readRDS(file.path(path.to.03.output, "cluster_markers.rds"))
}
```

## Heatmap: Top N up-gene in each clusters
```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=15, fig.height=10}
topN <- 10
tmp.cluster.markers.up <- subset(cluster.markers, cluster.markers$avg_log2FC >= 0 & cluster.markers$p_val_adj <= 0.05)

selected.genes <- data.frame()
for (cluster.id in unique(tmp.cluster.markers.up$cluster)){
  tmp <- subset(tmp.cluster.markers.up, tmp.cluster.markers.up$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  tmp.genes <- head(tmp, topN)
  selected.genes <- rbind(selected.genes, tmp.genes)
}

p <- DoHeatmap(object = s.obj, features = selected.genes$gene) + scale_fill_viridis_c(option = "plasma")
ggsave(plot = p, filename = "Heatmap_topN_up_genes_in_all_clusters.svg", path = file.path(path.to.save.figures), device = "svg", width = 28, height = 20, dpi = 300)
```


```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=28, fig.height=20}
print(p)
```


### Tables of all cluster marker genes {.tabset}
```{r echo=FALSE, include=FALSE}
cluster.markers %>% create_dt()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=30, fig.height=20}

for (cluster.id in sort(unique(cluster.markers$cluster))){
  tmp.cluster.markers <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  tmp.cluster.markers <- tmp.cluster.markers %>% mutate_if(is.numeric, round, 6) 
  
  cat(paste("\n\n#### Cluster: ", cluster.id, "##\n"))
  print( htmltools::tagList(datatable(tmp.cluster.markers, extensions = 'Buttons',
                filter = "top",
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All")),
                               columnDefs = list(list(
                                 targets = "_all",
                                 render = JS(
                                   "function(data, type, row, meta) {",
                                   "return type === 'display' && data != null && data.length > 100 ?",
                                   "'<span title=\"' + data + '\">' + data.substr(0, 100) + '...</span>' : data;",
                                   "}")
                               ))
                ))))
  cat("\n \n")
}

```


### Top-12 cluster marker genes `FeaturePlot` {.tabset}
Here we just show the top-12 genes (ranked by `avg_log2FC`) for each cluster. 

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
for (cluster.id in sort(unique(s.obj@meta.data$seurat_clusters))){
  DefaultAssay(s.obj) <- "RNA"
  cat(sprintf("#### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- FeaturePlot(object = s.obj, reduction = "INTE_UMAP", features = head(tmp.cluster.markers, 12)$gene, ncol = 3, label = TRUE, pt.size = 0.5, label.size = 5, label.color = "red")  
  print(p)
  cat("\n \n")
}
```

### Top-12 cluster marker genes `DotPlot` {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=30, fig.height=20}
for (cluster.id in sort(unique(s.obj@meta.data$seurat_clusters))){
  DefaultAssay(s.obj) <- "RNA"
  cat(sprintf("#### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- DotPlot(object = s.obj, features = head(tmp.cluster.markers, 9)$gene)  
  print(p)
  cat("\n \n")
}
```

### Top-12 cluster marker genes `Vlnplot` {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=30, fig.height=20}
for (cluster.id in sort(unique(s.obj@meta.data$seurat_clusters))){
  DefaultAssay(s.obj) <- "RNA"
  cat(sprintf("#### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- VlnPlot(object = s.obj, features = head(tmp.cluster.markers, 9)$gene)  
  print(p)
  cat("\n \n")
}
```

# Distribution of YFP cells in all two samples
## Sampling cells from *YFP- cells* and compare Shannon entropy with YFP+ cells
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=15, fig.height=10}
all.sampling.sizes <- c(100, 200, 300, 400, 500)

sampling.shannon.df <- data.frame()
for (sampling.size in all.sampling.sizes){
  all.sampling.shannon <- c()

  for (i in seq(1, 100)){
    sampling.cells <- sample(non.yfp.cells, sampling.size)
    sampling.shannon <- calculate_shannon_entropy(barcodes = sampling.cells, input.s.obj = s.obj)
    all.sampling.shannon <- c(all.sampling.shannon, sampling.shannon)
  }
  
  tmpdf <- data.frame(data = all.sampling.shannon)
  tmpdf$sampling_size <- sampling.size
  sampling.shannon.df <- rbind(sampling.shannon.df, tmpdf)
}

yfp.shannon.df <- data.frame(data = c(calculate_shannon_entropy(barcodes = yfp.cells, input.s.obj = s.obj)))
yfp.shannon.df$sampling_size <- "YFP.Shannon"
sampling.shannon.df <- rbind(sampling.shannon.df, yfp.shannon.df)
sampling.shannon.df$sampling_size <- factor(sampling.shannon.df$sampling_size, levels = c("YFP.Shannon", all.sampling.sizes))

p <- ggplot(data = sampling.shannon.df, aes(x = sampling_size, y = data, fill = sampling_size)) + geom_boxplot()
ggsave(plot = p, filename = "[validation]sampling_cells_Shannon_entropy_from_nonYFP_compare_YFP.svg", path = path.to.save.figures, device = "svg", dpi = 300, width = 15, height = 10)

writexl::write_xlsx(sampling.shannon.df, file.path(path.to.save.table, "[validation]sampling_cells_Shannon_entropy_from_nonYFP_compare_YFP.xlsx"))
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=15, fig.height=10}
print(p)
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=15, fig.height=10}
print(sprintf("Shannon entropy of YFP cells: %s", calculate_shannon_entropy(barcodes = yfp.cells, input.s.obj = s.obj)))
```

# VDJ data analysis
## All clones (in all 2 samples, both YFP+ and YFP-) and Shannon entropy (all colonizations/samples)

### Summary table

- `count`: Total number of cells. 

- `Shannon.entropy`: Shannon entropy of the clone, calculated in all clusters of all two samples in both condition YFP+ and YFP -

- `count.Sample132/133`: Number of cells in Sample 132/133.

- `count.in.YFP`: number of YFP+ cells of the clone.

- `count.in.nonYFP`: number of YFP- cells of the clone. 

- `clone.in.YFP`: "yes" if the clone also appear in YFP+ cells.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=15, fig.height=10}
meta.data <- s.obj@meta.data %>% rownames_to_column("barcode")
all.shannon.entropy <- s.obj@meta.data %>% subset(select = c(CTstrict)) %>% table() %>% as.data.frame() %>%
  subset(is.na(CTstrict) == FALSE) 
colnames(all.shannon.entropy) <- c("clone", "count")

all.shannon.entropy <- all.shannon.entropy %>%
  rowwise() %>% 
  mutate(Shannon.entropy = ifelse(count >= 5, calculate_shannon_entropy(input.s.obj =  s.obj, barcodes = subset(meta.data, meta.data$CTstrict == clone)$barcode), NA)) 

all.shannon.entropy <- all.shannon.entropy %>% rowwise() %>% 
  mutate(count.Sample132 = subset(s.obj@meta.data, s.obj@meta.data$CTstrict == clone & s.obj@meta.data$name == "Sample_132") %>% nrow()) %>%
  mutate(count.Sample133 = subset(s.obj@meta.data, s.obj@meta.data$CTstrict == clone & s.obj@meta.data$name == "Sample_133") %>% nrow()) %>%
  mutate(count.in.YFP = subset(s.obj.yfp@meta.data, s.obj.yfp@meta.data$CTstrict == clone) %>% nrow()) %>%
  mutate(count.in.nonYFP = subset(s.obj.nonyfp@meta.data, s.obj.nonyfp@meta.data$CTstrict == clone) %>% nrow()) %>%
  mutate(count.in.YFP.132 = subset(s.obj.yfp@meta.data, s.obj.yfp@meta.data$CTstrict == clone & s.obj.yfp@meta.data$name == "Sample_132") %>% nrow()) %>%
  mutate(count.in.YFP.133 = subset(s.obj.yfp@meta.data, s.obj.yfp@meta.data$CTstrict == clone & s.obj.yfp@meta.data$name == "Sample_133") %>% nrow()) %>%
  mutate(clone_in_yfp = ifelse(count.in.YFP != 0, "yes", "no"))
```


### Sampling cells and calculate Shannon entropy {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=15, fig.height=10}
for (num.sampling.cells in c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100)){
  all.cells <- colnames(s.obj)
  all.sampling.entropies <- c()
  
  if (file.exists(file.path(path.to.save.table, sprintf("[validation]clonal_Shannon_entropy_all_cells_sampling_%s.xlsx", num.sampling.cells))) == FALSE){
      for (count in seq(1, 1000)){
        sampling.cells <- sample(all.cells, num.sampling.cells)
        tmp.sampling.shannon.entropy <- calculate_shannon_entropy(sampling.cells, s.obj)
        all.sampling.entropies <- c(all.sampling.entropies, tmp.sampling.shannon.entropy)
        }
  
      samplingdf <- data.frame(data = all.sampling.entropies)
      colnames(samplingdf) <- c("Shannon.entropy")
      samplingdf$data <- to_vec( for (item in seq(1, 1000)) sprintf("sampling_%s", item))
      samplingdf <- samplingdf[c("data", "Shannon.entropy")]
      samplingdf$case <- "Sampling"
      
      tmp.all.shannon.entropy <- all.shannon.entropy %>% subset(select = c(clone, Shannon.entropy))
      colnames(tmp.all.shannon.entropy) <- c("data", "Shannon.entropy")
      tmp.all.shannon.entropy$case <- "real_clone"
      validationdf <- rbind(tmp.all.shannon.entropy, samplingdf)
      
      writexl::write_xlsx(validationdf, file.path(path.to.save.table, sprintf("[validation]clonal_Shannon_entropy_all_cells_sampling_%s.xlsx", num.sampling.cells)))
  } else {
    validationdf <- readxl::read_xlsx(file.path(path.to.save.table, sprintf("[validation]clonal_Shannon_entropy_all_cells_sampling_%s.xlsx", num.sampling.cells)))
  }
  p <- validationdf %>% ggplot(aes(x = case, y = Shannon.entropy)) + geom_boxplot() +
    ggtitle(sprintf("Shannon entropy: Real clone vs sampling %s cells", num.sampling.cells))
  cat(sprintf("#### Sampling %s cells \n", num.sampling.cells))
  print(p)
  cat("\n \n")
}
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=15, fig.height=10}
all.shannon.entropy %>% arrange(desc(count)) %>% create_dt()
```

## YFP clones only

### Summary table
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=15, fig.height=10}
meta.data <- s.obj.yfp@meta.data %>% rownames_to_column("barcode")
all.shannon.entropy.yfp <- s.obj.yfp@meta.data %>% subset(select = c(CTstrict)) %>% table() %>% as.data.frame() %>%
  subset(is.na(CTstrict) == FALSE) 
colnames(all.shannon.entropy.yfp) <- c("clone", "count")

all.shannon.entropy.yfp <- all.shannon.entropy.yfp %>%
  rowwise() %>% 
  mutate(Shannon.entropy = ifelse(count >= 5, calculate_shannon_entropy(input.s.obj =  s.obj.yfp, barcodes = subset(meta.data, meta.data$CTstrict == clone)$barcode), NA)) 

all.shannon.entropy.yfp %>% arrange(desc(count)) %>% create_dt()
```

### Sampling cells and calculate Shannon entropy {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=15, fig.height=10}
for (num.sampling.cells in c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100)){
  all.cells <- colnames(s.obj.yfp)
  all.sampling.entropies <- c()
  
  if (file.exists(file.path(path.to.save.table, sprintf("[validation]clonal_Shannon_entropy_YFP_cells_sampling_%s.xlsx", num.sampling.cells))) == FALSE){
    for (count in seq(1, 1000)){
      sampling.cells <- sample(all.cells, num.sampling.cells)
      tmp.sampling.shannon.entropy <- calculate_shannon_entropy(sampling.cells, s.obj.yfp)
      all.sampling.entropies <- c(all.sampling.entropies, tmp.sampling.shannon.entropy)
    }
    
    samplingdf <- data.frame(data = all.sampling.entropies)
    colnames(samplingdf) <- c("Shannon.entropy")
    samplingdf$data <- to_vec( for (item in seq(1, 1000)) sprintf("sampling_%s", item))
    samplingdf <- samplingdf[c("data", "Shannon.entropy")]
    samplingdf$case <- "Sampling"
    
    tmp.all.shannon.entropy <- all.shannon.entropy.yfp %>% subset(select = c(clone, Shannon.entropy))
    colnames(tmp.all.shannon.entropy) <- c("data", "Shannon.entropy")
    tmp.all.shannon.entropy$case <- "real_clone"
    validationdf <- rbind(tmp.all.shannon.entropy, samplingdf)
      
    writexl::write_xlsx(validationdf, file.path(path.to.save.table, sprintf("[validation]clonal_Shannon_entropy_YFP_cells_sampling_%s.xlsx", num.sampling.cells)))
  } else {
    validationdf <- readxl::read_xlsx(file.path(path.to.save.table, sprintf("[validation]clonal_Shannon_entropy_YFP_cells_sampling_%s.xlsx", num.sampling.cells)))
  }
  p <- validationdf %>% ggplot(aes(x = case, y = Shannon.entropy)) + geom_boxplot() +
    ggtitle(sprintf("Shannon entropy: Real clone vs sampling %s cells", num.sampling.cells))
  cat(sprintf("#### Sampling %s cells \n", num.sampling.cells))
  print(p)
  cat("\n \n")
}
```

## nonYFP clones only

### Summary table
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=15, fig.height=10}
meta.data <- s.obj.nonyfp@meta.data %>% rownames_to_column("barcode")
all.shannon.entropy.nonyfp <- s.obj.nonyfp@meta.data %>% subset(select = c(CTstrict)) %>% table() %>% as.data.frame() %>%
  subset(is.na(CTstrict) == FALSE) 
colnames(all.shannon.entropy.nonyfp) <- c("clone", "count")

all.shannon.entropy.nonyfp <- all.shannon.entropy.nonyfp %>%
  rowwise() %>% 
  mutate(Shannon.entropy = ifelse(count >= 5, calculate_shannon_entropy(input.s.obj =  s.obj.nonyfp, barcodes = subset(meta.data, meta.data$CTstrict == clone)$barcode), NA)) 

all.shannon.entropy.nonyfp %>% arrange(desc(count)) %>% create_dt()
```

### Sampling cells and calculate Shannon entropy {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=15, fig.height=10}
for (num.sampling.cells in c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100)){
  all.cells <- colnames(s.obj.nonyfp)
  all.sampling.entropies <- c()
  
  if (file.exists(file.path(path.to.save.table, sprintf("[validation]clonal_Shannon_entropy_nonYFP_cells_sampling_%s.xlsx", num.sampling.cells))) == FALSE){
    for (count in seq(1, 1000)){
      sampling.cells <- sample(all.cells, num.sampling.cells)
      tmp.sampling.shannon.entropy <- calculate_shannon_entropy(sampling.cells, s.obj.nonyfp)
      all.sampling.entropies <- c(all.sampling.entropies, tmp.sampling.shannon.entropy)
    }
    
    samplingdf <- data.frame(data = all.sampling.entropies)
    colnames(samplingdf) <- c("Shannon.entropy")
    samplingdf$data <- to_vec( for (item in seq(1, 1000)) sprintf("sampling_%s", item))
    samplingdf <- samplingdf[c("data", "Shannon.entropy")]
    samplingdf$case <- "Sampling"
    
    tmp.all.shannon.entropy <- all.shannon.entropy.nonyfp %>% subset(select = c(clone, Shannon.entropy))
    colnames(tmp.all.shannon.entropy) <- c("data", "Shannon.entropy")
    tmp.all.shannon.entropy$case <- "real_clone"
    validationdf <- rbind(tmp.all.shannon.entropy, samplingdf)
      
    writexl::write_xlsx(validationdf, file.path(path.to.save.table, sprintf("[validation]clonal_Shannon_entropy_nonYFP_cells_sampling_%s.xlsx", num.sampling.cells)))
  } else {
    validationdf <- readxl::read_xlsx(file.path(path.to.save.table, sprintf("[validation]clonal_Shannon_entropy_nonYFP_cells_sampling_%s.xlsx", num.sampling.cells)))
  }
  p <- validationdf %>% ggplot(aes(x = case, y = Shannon.entropy)) + geom_boxplot() +
    ggtitle(sprintf("Shannon entropy: Real clone vs sampling %s cells", num.sampling.cells))
  cat(sprintf("#### Sampling %s cells \n", num.sampling.cells))
  print(p)
  cat("\n \n")
}
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=15, fig.height=10}
```

