---
title: "Monocle2 trajectory analysis"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    number_sections: true
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    theme: lumen
params:
  dataset.name: NA
---
  
```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
$(document).ready(function() {
  $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
  // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

<style type="text/css">
    div.datatables { height: auto !important;}
</style>

```{r echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, fig.height=10, fig.width=14, results='hide'}
my_random_seed <- 42
set.seed(my_random_seed)

scrna_pipeline_src <- "/home/hieunguyen/CRC1382/src_2023/src_pipeline/scRNA_GEX_pipeline/processes_src"
source(file.path(scrna_pipeline_src, "import_libraries.R"))
source(file.path(scrna_pipeline_src, "helper_functions.R"))

path.to.project.src <- "/home/hieunguyen/CRC1382/src_2023/FHager_datasets/NEW_20240508"
source(file.path(path.to.project.src, "helper_functions.R"))
source(file.path(path.to.project.src, "config.R"))

#####-----------------------------------------------------------------------#####
##### install monocle
#####-----------------------------------------------------------------------#####
# if ("monocle3" %in% installed.packages()){
# remove.packages("monocle3")
# devtools::install_github("cysouw/qlcMatrix")
# install.packages("DDRTree")
# install.packages("densityClust")
# BiocManager::install("monocle.objSingleCell", update = FALSE)
# install.packages("fastICA")
# BiocManager::install("biocViews", update = FALSE)
# # remove.packages("BiocGenerics")
# BiocManager::install("HSMMSingleCell", update = FALSE)
# # install.packages("/media/hieunguyen/HD0/storage/offline_pkgs/monocle_2.30.0.tar.gz", type = "source", repos = NULL)
# install.packages("/media/hieunguyen/HD0/storage/offline_pkgs/monocle_2.30.0/monocle", type = "source", repos = NULL)
# BiocManager::install("tradeSeq", update = FALSE)
# }
#####-----------------------------------------------------------------------#####

if ("svglite" %in% installed.packages() == FALSE){
  install.packages("svglite")
}
library(devtools)
library(monocle)

# dataset.name <- params$dataset.name
dataset.name <- "2nd_dataset_removed_5_6.without_reInt.res1"
# dataset.name <- "1st_dataset_removed_7_9_removed_16_removed_9.with_reInt.res1"
  
outdir <- "/media/hieunguyen/HD0/outdir/CRC1382"

path.to.output <- file.path(outdir, "BSimons", "OUTPUT", "THESIS_OUTPUT_20231026")
path.to.01.output <- file.path(path.to.output, "01_output", dataset.name)
path.to.04.output <- file.path(path.to.output, "04_output", dataset.name)
dir.create(path.to.04.output, showWarnings = FALSE, recursive = TRUE)

s.obj <- readRDS(file.path(path.to.01.output, sprintf("%s.rds", dataset.name)))

if (file.exists(file.path(path.to.04.output, "monocle_obj.rds")) == FALSE){
  monocle.obj <- run_monocle2(s.obj, path.to.04.output)
  write.csv(data.frame(status = c("finished saving monocle obj")), file.path(path.to.04.output, "finished.csv"))
} else {
  print("monocle object exists!")
  monocle.obj <- readRDS(file.path(path.to.04.output, "monocle_obj.rds")) 
}




```


# Plot
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE)
```

## UMAP
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
plot_cell_trajectory(monocle.obj, color_by = "seurat_clusters")
```

## Cell trajectory from `monocle2`
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
plot_cell_trajectory(monocle.obj, show_branch_points = TRUE, show_tree = TRUE)
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
plot_cell_trajectory(monocle.obj, color_by = "Pseudotime")
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
##### Add pseudotime information to the main seurat object
seurat.metadata <- s.obj@meta.data %>% rownames_to_column("barcode")
monocle.metadata <- pData(monocle.obj) %>% rownames_to_column("barcode") %>%
  subset(select = c(barcode, Pseudotime, State))

meta.data <- merge(seurat.metadata, monocle.metadata, by.x = "barcode", by.y = "barcode") %>%
  column_to_rownames("barcode")
meta.data <- meta.data[row.names(s.obj@meta.data), ]
s.obj <- AddMetaData(object = s.obj, metadata = meta.data$Pseudotime, col.name = "Pseudotime")
s.obj <- AddMetaData(object = s.obj, metadata = meta.data$State, col.name = "State")
```

## UMAP: grouped by trajectory STATE
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, group.by = "State")
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
FeaturePlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, features = c("Pseudotime"))
```


## Differential gene expression test along trajectories (VGAM)
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
if (file.exists(file.path(path.to.04.output, "pseudotime.markers.rds")) == FALSE){
  pseudotime.markers <- differentialGeneTest(monocle.obj, fullModelFormulaStr = "~sm.ns(Pseudotime)", cores = 16, verbose = TRUE)
  saveRDS(pseudotime.markers, file.path(path.to.04.output, "pseudotime.markers.rds"))
} else {
  pseudotime.markers <- readRDS(file.path(path.to.04.output, "pseudotime.markers.rds"))
}

```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
sig_gene_names <- row.names(subset(pseudotime.markers, qval < 0.1))
plot_pseudotime_heatmap(monocle.obj[head(sig_gene_names, 10),],
                num_clusters = 3,
                cores = 1,
                show_rownames = T)
```

# Differential gene expression test (TradeSeq)
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
# devtools::install_github("statOmics/tradeSeq")
library(tradeSeq)

info <- extract_monocle_info(monocle.obj)
cellWeights <- info$cellWeights
pseudotime <- info$pseudotime

meta.data <- s.obj@meta.data %>% rownames_to_column("barcode")
tmp.cellWeights <- data.frame(cellWeights) %>% rownames_to_column("barcode")
meta.data <- merge(meta.data, tmp.cellWeights, by.x = "barcode", by.y = "barcode")
meta.data <- meta.data %>% column_to_rownames("barcode")
meta.data <- meta.data[row.names(s.obj@meta.data), ]

for (tmp.col in colnames(cellWeights)){
  s.obj <- AddMetaData(object = s.obj, metadata = meta.data[[tmp.col]], col.name = tmp.col)
}
```

A matrix of cell weights defining the probability that a cell belongs to a particular lineage. Each row represents a cell and each column represents a lineage. If only a single lineage, provide a matrix with one column containing all values of 1.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
for (tmp.col in colnames(cellWeights)){
  p <- DimPlot(object = s.obj, reduction = "INTE_UMAP", group.by = tmp.col, label = TRUE, pt.size = 1) + scale_color_manual(values = c("gray", "red"))
  print(p)
}
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
print(sprintf("Number of lineage in this dataset: %s", ncol(cellWeights)))
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
if (file.exists(file.path(path.to.04.output, "sce.rds")) == FALSE){
  sce <- fitGAM(counts = Biobase::exprs(monocle.obj),
              cellWeights = cellWeights,
              pseudotime = pseudotime, 
              sce = TRUE, 
              nknots = 6,
              parallel = TRUE, 
              genes = intersect(row.names(Biobase::exprs(monocle.obj)), VariableFeatures(s.obj)))
  saveRDS(sce, file.path(path.to.04.output, "sce.rds"))  
} else {
  sce <- readRDS(file.path(path.to.04.output, "sce.rds"))
}
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
#####----------------------------------------------------------------------#####
##### Test within lineage
#####----------------------------------------------------------------------#####
if (file.exists(file.path(path.to.04.output, "tradeseq_test_res.rds")) == FALSE){
  tradeseq.testres <- hash()
  ##### Association test
  tmp <- associationTest(sce)
  tmp <- subset(tmp, is.na(tmp$pvalue) == FALSE)
  tmp$p.adjust <- p.adjust(tmp$pvalue, method = "BH")
  tmp <- subset(tmp, tmp$p.adjust <= 0.05) %>% arrange(desc(meanLogFC)) %>% 
    rownames_to_column("Gene")
  tradeseq.testres[["Association_test"]] <- tmp
  
  ##### Start vs End test
  tmp <- startVsEndTest(sce)
  tmp <- subset(tmp, is.na(tmp$pvalue) == FALSE)
  tmp$p.adjust <- p.adjust(tmp$pvalue, method = "BH")
  tmp <- subset(tmp, tmp$p.adjust <= 0.05)  %>% 
    rownames_to_column("Gene")
  tradeseq.testres[["StartVsEnd_test"]] <- tmp
  
  #####----------------------------------------------------------------------#####
  ##### Test across lineage
  #####----------------------------------------------------------------------#####
  if (ncol(cellWeights) >= 2){
      ##### End vs End test
    tmp <- diffEndTest(sce)
    tmp <- subset(tmp, is.na(tmp$pvalue) == FALSE)
    tmp$p.adjust <- p.adjust(tmp$pvalue, method = "BH")
    tmp <- subset(tmp, tmp$p.adjust <= 0.05) %>%
      rownames_to_column("Gene")
    tradeseq.testres[["diff_End_test"]] <- tmp
    
    ##### Pattern test
    tmp <- patternTest(sce)
    tmp <- subset(tmp, is.na(tmp$pvalue) == FALSE)
    tmp$p.adjust <- p.adjust(tmp$pvalue, method = "BH")
    tmp <- subset(tmp, tmp$p.adjust <= 0.05) %>% 
      rownames_to_column("Gene")
    tradeseq.testres[["Pattern_test"]] <- tmp
    
    ##### early DE test
    tmp <- earlyDETest(sce,)
    tmp <- subset(tmp, is.na(tmp$pvalue) == FALSE)
    tmp$p.adjust <- p.adjust(tmp$pvalue, method = "BH")
    tmp <- subset(tmp, tmp$p.adjust <= 0.05) %>% 
      rownames_to_column("Gene")
    tradeseq.testres[["EarlyDE_test"]] <- tmp
  }
  
  saveRDS(tradeseq.testres, file.path(path.to.04.output, "tradeseq_test_res.rds"))
} else {
  tradeseq.testres <- readRDS(file.path(path.to.04.output, "tradeseq_test_res.rds"))
}
```

## All test results (table) {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, include=FALSE}
tradeseq.testres$Association_test %>% create_dt()
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
for (input.test in names(tradeseq.testres)){
  cat(sprintf("### Test: %s \n", input.test))
  tmp <- tradeseq.testres[[input.test]]
  print( htmltools::tagList(datatable(tmp, extensions = 'Buttons',
                filter = "top",
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All")),
                               columnDefs = list(list(
                                 targets = "_all",
                                 render = JS(
                                   "function(data, type, row, meta) {",
                                   "return type === 'display' && data != null && data.length > 100 ?",
                                   "'<span title=\"' + data + '\">' + data.substr(0, 100) + '...</span>' : data;",
                                   "}")
                               ))
                ))))
  cat("\n \n")
}
```

# Branch analysis (monocle-BEAM)
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
all.branch.points <- monocle.obj@auxOrderingData$DDRTree$branch_points
if (length(all.branch.points) >= 1){
  names(all.branch.points) <- seq(1, length(all.branch.points))

  all.principal.points <- data.frame(monocle.obj@reducedDimK) %>% t() %>% as.data.frame()
  colnames(all.principal.points) <- c("V1", "V2")
  all.principal.points <- all.principal.points %>% rownames_to_column("principal_point") %>%
    rowwise() %>% 
    mutate(type = ifelse(principal_point %in% all.branch.points, "branch", "normal")) %>%
    mutate(branch_point = ifelse(principal_point %in% all.branch.points, principal_point, NA))
  
  data.frame(all.branch.points) %>% rownames_to_column("index") %>% create_dt()  
} else {
  print("No branch found")
}

```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
if (length(all.branch.points) >= 1){
  all.principal.points %>% ggplot(aes(x = V1, y = V2, color = type, label = branch_point)) + geom_point() + scale_color_manual(values = c("red", "gray")) + geom_text_repel()  
}

```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
if (length(all.branch.points) >= 1){
  all.BEAM_res <- hash()
  for (i in seq(1, length(all.branch.points))){
    if (file.exists(file.path(path.to.04.output, sprintf("BEAM_res.branch_point_%s.rds", i))) == FALSE){
      BEAM_res <- BEAM(monocle.obj, branch_point = i, cores = 10)
      BEAM_res <- BEAM_res[order(BEAM_res$qval),]
      BEAM_res <- BEAM_res[,c("gene_short_name", "pval", "qval")]
      saveRDS(BEAM_res, file.path(path.to.04.output, sprintf("BEAM_res.branch_point_%s.rds", i)))  
    } else {
      BEAM_res <- readRDS(file.path(path.to.04.output, sprintf("BEAM_res.branch_point_%s.rds", i)))
    }
    all.BEAM_res[[sprintf("p%s", i)]] <- BEAM_res[order(BEAM_res$qval),]  
  }  
} else {
  print("NO branch point to conduct BEAM analysis")
}

```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
if (length(all.branch.points) >= 1){
  tmp.plot <- list()
  for (i in names(all.BEAM_res)){
    BEAM_res <- all.BEAM_res[[i]]
    selected.BEAM.genes <-  head(BEAM_res, 100)
    p <- plot_genes_branched_heatmap(monocle.obj[selected.BEAM.genes$gene_short_name,],
                                     branch_point = 1,
                                     num_clusters = 4,
                                     cores = 1,
                                     use_gene_short_name = T,
                                     show_rownames = T, 
                                     return_heatmap = TRUE)
    tmp.plot[[i]] <- ggplotify::as.ggplot(p$ph_res)
    dev.off()
    # ggsave(plot = p$ph_res, filename = "BEAM_branch_DE_genes_top100.svg", device = "svg", path = file.path(path.to.04.output), dpi = 300, width = 20, height = 15)
  }  
}
```

## BEAM results {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
if (length(all.branch.points) >= 1){
  for (i in names(tmp.plot)){
    cat(sprintf("### Branch point %s, %s \n", i, all.branch.points[[as.numeric(str_replace(i, "p", ""))]]))
    print(tmp.plot[[i]])
    cat("\n \n")
  }
} else {
  print("NO BRANCH POINT TO CONDUCT BEAM ANALYSIS")
}

```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
```

