---
title: "Integration and Downstream analysis, Quantile: `r params$chosen.quantile`, Integration: `r params$integration.case`"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: no
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: true
    theme: lumen
params:
  outdir: NA
  PROJECT: NA
  config.version: NA
  output.version: NA
  chosen.quantile: NA
  integration.case: NA
---


```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

<style type="text/css">
    div.datatables { height: auto !important;}
</style>


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=14, fig.height=10}
#####----------------------------------------------------------------------#####
#
# trnguyen@ukaachen.de
#
#####----------------------------------------------------------------------#####
# gc()
# rm(list = ls())

scrna_pipeline_src <- "/home/hieunguyen/CRC1382/src_2023/src_pipeline/scRNA_GEX_pipeline/processes_src"
source(file.path(scrna_pipeline_src, "import_libraries.R"))
source(file.path(scrna_pipeline_src, "helper_functions.R"))
source(file.path(scrna_pipeline_src, "s8_integration_and_clustering.R"))

path.to.project.src <- "/home/hieunguyen/CRC1382/src_2023/bcr_data_analysis/240805_data_analysis"
source(file.path(path.to.project.src, "config.R"))
#####----------------------------------------------------------------------#####
# CONFIGURATIONS 
#####----------------------------------------------------------------------#####
analysis.round <- "1st"
chosen.seed <- 42
num.dim.integration <- 25 
num.PCA <- 25
num.dim.cluster <- 25
num.PC.used.in.Clustering <- 25
num.PC.used.in.UMAP <- 25
my_random_seed <- 42


all.samples <- c("M1", "M2", "M3", "P1", "P2", "P3")
all.quantiles <- c(0.99, 0.95, 0.90, 0.85)
all.integration.case <- list(
  all_samples = all.samples,
  mouse1 = c("M1", "P1"),
  mouse2 = c("M2", "P2"),
  mouse3 = c("M3", "P3")
)

#####----------------------------------------------------------------------#####
##### input arguments
#####----------------------------------------------------------------------#####

# outdir <- "/home/hieunguyen/CRC1382/outdir"
# PROJECT <- "240805_BSimons"
# output.version <- "20240820"
# config.version <- "default"
# chosen.quantile <- 0.99
# integration.case <- "all_samples"

outdir <- params$outdir
PROJECT <- params$PROJECT
output.version <- params$output.version
config.version <- params$config.version
chosen.quantile <- params$chosen.quantile
integration.case <- params$integration.case

path.to.main.input <- file.path(outdir, PROJECT, output.version, config.version)
path.to.main.output <- file.path(outdir, PROJECT, "data_analysis", output.version, config.version)

path.to.01.output <- file.path(path.to.main.output, "01_output")
path.to.04.output <- file.path(path.to.main.output, "04_output", output.version, config.version, sprintf("quantile_%s", chosen.quantile))
path.to.05.output <- file.path(path.to.main.output, "05_output", output.version, config.version, sprintf("quantile_%s", chosen.quantile), integration.case)
dir.create(path.to.05.output, showWarnings = FALSE, recursive = TRUE)

path.to.input.s.obj <- file.path(path.to.04.output, integration.case, "s8_output", sprintf("%s.output.s8.rds", PROJECT))
s.obj <- readRDS(path.to.input.s.obj)
```

# UMAP of all samples {.tabset}

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
for (sample.id in unique(s.obj$name)){
  cat(sprintf("## Sample: %s \n", sample.id))
  p <- DimPlot(object = subset(s.obj, name == sample.id), reduction = "RNA_UMAP", label = TRUE, label.box = TRUE, label.size = 10, repel = TRUE) + 
  ggtitle(sprintf("UMAP Sample %s: All clusters", sample.id)) +
  theme(axis.text = element_text(size=20, face = "bold"),
        axis.title = element_text(size=20, face = "bold"), 
        title = element_text(size = 20, face = "bold"))
  print(p)
  cat("\n \n")
}

```

## Integrate all samples and show them on UMAP

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
DimPlot(object = s.obj, reduction = "RNA_UMAP", label = TRUE, group.by = "name", label.box = TRUE, label.size = 10, repel = TRUE) +
  ggtitle("UMAP: all samples before performing data integration") + 
  theme(axis.text = element_text(size=20, face = "bold"),
        axis.title = element_text(size=20, face = "bold"), 
        title = element_text(size = 20, face = "bold"))
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, group.by = "name", label.box = TRUE, label.size = 10, repel = TRUE) +
  ggtitle("UMAP: all samples after performing data integration") + 
  theme(axis.text = element_text(size=20, face = "bold"),
        axis.title = element_text(size=20, face = "bold"), 
        title = element_text(size = 20, face = "bold"))
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=28, fig.height=20}
DimPlot(object = s.obj, reduction = "INTE_UMAP", label = FALSE, split.by = "name", ncol = 3, label.box = TRUE, label.size = 10, repel = TRUE) +
  ggtitle("UMAP: all samples after performing data integration") + 
  theme(axis.text = element_text(size=20, face = "bold"),
        axis.title = element_text(size=20, face = "bold"), 
        title = element_text(size = 20, face = "bold"))
```

## Clustering after integration
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=28, fig.height=20}
DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, label.size = 10, repel = TRUE) +
  ggtitle("UMAP: all samples after performing data integration") + 
  theme(axis.text = element_text(size=20, face = "bold"),
        axis.title = element_text(size=20, face = "bold"), 
        title = element_text(size = 20, face = "bold"))
```


# Cluster markers of integrated dataset
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}

if (file.exists(file.path(path.to.05.output, "integrated_cluster_markers.rds")) == FALSE){
  integrated.cluster.markers <- FindAllMarkers(object = s.obj, assay = "RNA", test.use = "wilcox")
  saveRDS(integrated.cluster.markers, file.path(path.to.05.output, "integrated_cluster_markers.rds"))   
} else {
  integrated.cluster.markers <- readRDS(file.path(path.to.05.output, "integrated_cluster_markers.rds"))
}

integrated.cluster.markers <- subset(integrated.cluster.markers, integrated.cluster.markers$p_val_adj <= 0.05 & integrated.cluster.markers$avg_log2FC > 0)
```

## Top-12 cluster marker genes `FeaturePlot` {.tabset}
Here we just show the top-12 genes (ranked by `avg_log2FC`) for each cluster. 

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=28, fig.height=20}
for (cluster.id in sort(unique(s.obj@meta.data$seurat_clusters))){
  DefaultAssay(s.obj) <- "RNA"
  cat(sprintf("### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(integrated.cluster.markers, integrated.cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- FeaturePlot(object = s.obj, reduction = "INTE_UMAP", features = head(tmp.cluster.markers, 9)$gene, ncol = 3, label = TRUE, pt.size = 0.5, label.size = 5, label.color = "red")  
  print(p)
  cat("\n \n")
}
```


## Top-12 dot plot {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=28, fig.height=20}
for (cluster.id in sort(unique(s.obj@meta.data$seurat_clusters))){
  DefaultAssay(s.obj) <- "RNA"
  cat(sprintf("### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(integrated.cluster.markers, integrated.cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- DotPlot(object = s.obj, features = head(tmp.cluster.markers, 9)$gene)  
  print(p)
  cat("\n \n")
}
```

## Top-12 Violin plot {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=28, fig.height=20}
for (cluster.id in sort(unique(s.obj@meta.data$seurat_clusters))){
  DefaultAssay(s.obj) <- "RNA"
  cat(sprintf("### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(integrated.cluster.markers, integrated.cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- VlnPlot(object = s.obj, features = head(tmp.cluster.markers, 9)$gene)  
  print(p)
  cat("\n \n")
}
```

## Tables of all cluster marker genes {.tabset}
```{r echo=FALSE, include=FALSE}
integrated.cluster.markers %>% create_dt()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}

for (cluster.id in sort(unique(integrated.cluster.markers$cluster))){
  tmp.integrated.cluster.markers <- subset(integrated.cluster.markers, integrated.cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  tmp.integrated.cluster.markers <- tmp.integrated.cluster.markers %>% mutate_if(is.numeric, round, 6) 
  
  cat(paste("\n\n### Cluster: ", cluster.id, "##\n"))
  print( htmltools::tagList(datatable(tmp.integrated.cluster.markers, extensions = 'Buttons',
                filter = "top",
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All")),
                               columnDefs = list(list(
                                 targets = "_all",
                                 render = JS(
                                   "function(data, type, row, meta) {",
                                   "return type === 'display' && data != null && data.length > 100 ?",
                                   "'<span title=\"' + data + '\">' + data.substr(0, 100) + '...</span>' : data;",
                                   "}")
                               ))
                ))))
  cat("\n \n")
}

```




