---
title: "`r params$PROJECT` Quality Control Hashtag-Antibody 1, 2, 3. Different quantiles comparison (Sample: `r params$sample.id`)"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: no
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: true
    theme: lumen
params:
  PROJECT: NA
  sample.id: NA
  outdir: NA
  output.version: NA
  config.version: NA
---
  
```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
$(document).ready(function() {
  $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
  // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=20, fig.height=12}
scrna_pipeline_src <- "/home/hieunguyen/CRC1382/src_2023/src_pipeline/scRNA_GEX_pipeline/processes_src"
source(file.path(scrna_pipeline_src, "import_libraries.R"))
source(file.path(scrna_pipeline_src, "helper_functions.R"))

path.to.project.src <- "/home/hieunguyen/CRC1382/src_2023/bcr_data_analysis/240805_data_analysis"
source(file.path(path.to.project.src, "hashtag_notes.R"))

PROJECT <- params$PROJECT
sample.id <- params$sample.id
outdir <- params$outdir
output.version <- params$output.version
config.version <- params$config.version

# outdir <- "/home/hieunguyen/CRC1382/outdir"
# PROJECT <- "240805_BSimons"
# sample.id <- "M1"
# output.version <- "20240820"
# config.version <- "default"
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
print(sprintf("Data saved at %s", outdir))
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=20, fig.height=12}
path.to.main.input <- file.path(outdir, PROJECT, output.version, config.version, "1st_round")
path.to.main.output <- file.path(outdir, PROJECT, "data_analysis", output.version, config.version)
path.to.01.output <- file.path(path.to.main.output, "01_output")
dir.create(path.to.01.output, showWarnings = FALSE, recursive = TRUE)

path.to.before.filter.sobj <- file.path(path.to.main.input, sprintf("%s_1st_round/s1_output", sample.id), sprintf("%s.output.s1.rds", PROJECT))
before.filter.s.obj <- readRDS(path.to.before.filter.sobj)
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
DefaultAssay(before.filter.s.obj) <- "ADT"

all.hts <- num.hashtags[[sample.id]]
counts.adt <- GetAssayData(before.filter.s.obj, slot = "counts", assay = "ADT")
s.obj.HTO <- CreateSeuratObject(counts = counts.adt[all.hts, ] + 1, assay = "HTO") #### add pseudo count + 1 to avoide error "Cells with zero counts exist as a cluster"
s.obj.HTO <- NormalizeData(s.obj.HTO, normalization.method = "CLR")

maindf <- counts.adt[all.hts, ] %>% as.data.frame() %>% t() %>% as.data.frame() %>% rownames_to_column("barcode")
maindf.pivot <- maindf %>% pivot_longer(!barcode, names_to = "hashtag", values_to = "count")
```
# Boxplot: Distribution of "counts" in each hashtag

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
maindf.pivot %>% ggplot(aes(x = hashtag, y = count, fill = hashtag)) + geom_boxplot()
```

Note: In this analysis, we keep all cells obtained from the output of `CellRanger`; conditions on the percentage of Mitochondrial and Ambient RNA contamination have not been applied yet. 

# Quantile 0.99
## Cell hashtags demultiplexing

### Assigned cut-off 
```{r echo=FALSE}
chosen.quantile <- 0.99
s.obj.HTO <- HTODemux(s.obj.HTO, assay = "HTO", positive.quantile = chosen.quantile)
```

```{r echo=FALSE}
write.csv(s.obj.HTO@meta.data %>% rownames_to_column("barcode"), file.path(path.to.01.output, sprintf("metadata_HTO_%s_quantile_%s.csv", sample.id, chosen.quantile)))
```

### Summary table: Number of cells in each classified hashtag group

We perform a k-medoid clustering on the normalized Hashtag values, which initially separates cells into K(# of samples)+1 clusters. For each hashtag, we use the cluster with the lowest average value as the negative group. (Reference: https://satijalab.org/seurat/articles/hashing_vignette.html#8-hto-dataset-from-human-pbmcs)

- `Negative`: Cells that are classified as `negative` in all 3 hashtags.

- `Doublet`: Cells that are expressed in more than 1 hashtag based on the assigned cut-off.

- `Singlet`: Cells that are expressed in only 1 hashtag based on the assigned cut-off.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
sum.table <- table(s.obj.HTO$HTO_classification.global) %>% as.data.frame()
colnames(sum.table) <- c("assigned_group", "# cells")
sum.table %>% create_dt()

sum.table.099 <- sum.table
```

# Quantile 0.95
## Cell hashtags demultiplexing

### Assigned cut-off 
```{r echo=FALSE}
chosen.quantile <- 0.95
s.obj.HTO <- HTODemux(s.obj.HTO, assay = "HTO", positive.quantile = chosen.quantile)
```

```{r echo=FALSE}
write.csv(s.obj.HTO@meta.data %>% rownames_to_column("barcode"), file.path(path.to.01.output, sprintf("metadata_HTO_%s_quantile_%s.csv", sample.id, chosen.quantile)))
```

### Summary table: Number of cells in each classified hashtag group

We perform a k-medoid clustering on the normalized Hashtag values, which initially separates cells into K(# of samples)+1 clusters. For each hashtag, we use the cluster with the lowest average value as the negative group. (Reference: https://satijalab.org/seurat/articles/hashing_vignette.html#8-hto-dataset-from-human-pbmcs)

- `Negative`: Cells that are classified as `negative` in all 3 hashtags.

- `Doublet`: Cells that are expressed in more than 1 hashtag based on the assigned cut-off.

- `Singlet`: Cells that are expressed in only 1 hashtag based on the assigned cut-off.



```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
sum.table <- table(s.obj.HTO$HTO_classification.global) %>% as.data.frame()
colnames(sum.table) <- c("assigned_group", "# cells")
sum.table %>% create_dt()

sum.table.095 <- sum.table
```


# Quantile 0.90
## Cell hashtags demultiplexing

### Assigned cut-off 
```{r echo=FALSE}
chosen.quantile <- 0.90
s.obj.HTO <- HTODemux(s.obj.HTO, assay = "HTO", positive.quantile = chosen.quantile)
```

```{r echo=FALSE}
write.csv(s.obj.HTO@meta.data %>% rownames_to_column("barcode"), file.path(path.to.01.output, sprintf("metadata_HTO_%s_quantile_%s.csv", sample.id, chosen.quantile)))
```

### Summary table: Number of cells in each classified hashtag group

We perform a k-medoid clustering on the normalized Hashtag values, which initially separates cells into K(# of samples)+1 clusters. For each hashtag, we use the cluster with the lowest average value as the negative group. (Reference: https://satijalab.org/seurat/articles/hashing_vignette.html#8-hto-dataset-from-human-pbmcs)

- `Negative`: Cells that are classified as `negative` in all 3 hashtags.

- `Doublet`: Cells that are expressed in more than 1 hashtag based on the assigned cut-off.

- `Singlet`: Cells that are expressed in only 1 hashtag based on the assigned cut-off.



```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
sum.table <- table(s.obj.HTO$HTO_classification.global) %>% as.data.frame()
colnames(sum.table) <- c("assigned_group", "# cells")
sum.table %>% create_dt()

sum.table.090 <- sum.table
```


# Quantile 0.85
## Cell hashtags demultiplexing

### Assigned cut-off 
```{r echo=FALSE}
chosen.quantile <- 0.85
s.obj.HTO <- HTODemux(s.obj.HTO, assay = "HTO", positive.quantile = chosen.quantile)
```

```{r echo=FALSE}
write.csv(s.obj.HTO@meta.data %>% rownames_to_column("barcode"), file.path(path.to.01.output, sprintf("metadata_HTO_%s_quantile_%s.csv", sample.id, chosen.quantile)))
```

### Summary table: Number of cells in each classified hashtag group

We perform a k-medoid clustering on the normalized Hashtag values, which initially separates cells into K(# of samples)+1 clusters. For each hashtag, we use the cluster with the lowest average value as the negative group. (Reference: https://satijalab.org/seurat/articles/hashing_vignette.html#8-hto-dataset-from-human-pbmcs)

- `Negative`: Cells that are classified as `negative` in all 3 hashtags.

- `Doublet`: Cells that are expressed in more than 1 hashtag based on the assigned cut-off.

- `Singlet`: Cells that are expressed in only 1 hashtag based on the assigned cut-off.



```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
sum.table <- table(s.obj.HTO$HTO_classification.global) %>% as.data.frame()
colnames(sum.table) <- c("assigned_group", "# cells")
sum.table %>% create_dt()
sum.table.085 <- sum.table
```


# Conclusion
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
final.sum <- merge(sum.table.099, sum.table.095, by.x = "assigned_group", by.y = "assigned_group")
colnames(final.sum) <- c("assigned_group", "qt0.99", "qt0.95")
final.sum <- merge(final.sum, sum.table.090, by.x = "assigned_group", by.y = "assigned_group")
colnames(final.sum) <- c("assigned_group", "qt0.99", "qt0.95", "qt0.90")
final.sum <- merge(final.sum, sum.table.085, by.x = "assigned_group", by.y = "assigned_group")
colnames(final.sum) <- c("assigned_group", "qt0.99", "qt0.95", "qt0.90", "qt0.85")

final.sum %>% create_dt()
```
